---
title: "Working with Image Data"
author: "Christine Yiwen Yeh"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set-up: Loading/Installing Packages

First, let's get some practice installing and loading packages: 

```{r inst_pkgs, eval=FALSE, warning=FALSE, include=TRUE}
# from CRAN 
install.packages(c("ggplot2", "tibble", "devtools"))

# from Bioconductor 
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.14")
BiocManager::install("EBImage")

# from Github 
devtools::install_github("https://github.com/livnatje/DIALOGUE")
devtools::install_github("https://github.com/Jerby-Lab/opipes")
```

To load packages, call `library(YOURPACKAGENAME)`

```{r ld_pkgs, echo=TRUE, message=FALSE, warning=FALSE}
# from CRAN
library(ggplot2) 
library(tibble) 
library(pbapply)

# from Bioconductor
library(EBImage)

# from Github 
library(opipes) 
library(DIALOGUE)
```

## Let's work with some real data: 

For the purposes of this tutorial, we will load one sample from the ovarian 
cancer project: TMA-13, FOV-1 

```{r composite}
PATH <- "/Volumes/ljerby/HGSC_Profiling/Data/SMI/5 Raw data/SMI-0028_TMA_13/CellComposite/CellComposite_F001.jpg"
comp <- readImage(PATH)
display(comp, method = "raster")
```

_Wait, what did you just do?_

We just ran two functions from `EBImage`. A neat trick for helping you fetch the
documentation on any function, is by putting a `?` in front of the function name

```{r q_image_1,  eval=FALSE, warning=FALSE, include=TRUE}
?readImage 
```

_What's in an `Image` object in R?_

```{r q_image_2}
comp
```

## Reading in IF layers for processing

One of the things I'm working on whilst pre-processing the data, is to extract a 
quantitative measure of the IF intensity. These images are LARGE. Even loading 
them will take a bit. 

From the documentation we obtained from `Nanostring`, we know that each of the 
frames in the raw images corresponds to a particular immunofluorescent stain. 

```{r if_image_load}
PATH <- "/Volumes/ljerby/HGSC_Profiling/Data/SMI/5 Raw data/SMI-0028_TMA_13/RawMorphologyImages/20220222_094630_S2_C902_P99_N99_F001_Z005.TIF"
img <- readImage(PATH)
display(img, all = T)
```

## Exploring some outputs of Cell-Segmentation

In order to extract IF intensities, you want to be able to assign each pixel to
a cell. Lucky for me, Livnat has run cell-segmentation on our data. The `DAPI` 
(frame 5) and `Membrane (CD298/B2M)` was used to do the cell-segmentation.

```{r cell_seg_1}
# Reading and cleaning matrix names 
cellseg <- read.csv("/Volumes/ljerby/HGSC_Profiling/Results/Segmentation/SMI/Seg/SMI_T13_F001_whole-cell.csv")
colnames(cellseg) <- unlist(lapply(colnames(cellseg), function(x) {strsplit(x , split = "X0.")[[1]][2]}))
colnames(cellseg)[1] <- "0"
cellseg_raw <- cellseg
cellseg_raw[1000:1020, 1000:1009] 
```

Plotting the cell segmentation map

```{r cell_seg_2}
# Plot the cell segmentation map
cellseg[cellseg == 1001] <- 0.9 # if a pixel is part of the cell, set the image intensity to 0.9
display(1- Image(t(as.matrix(cellseg))),  method = "raster")
```

Reading and plotting the nuclear-segmentation only map: 
```{r cell_seg_3}
# Reading and cleaning matrix names 
nuclearseg <- read.csv("/Volumes/ljerby/HGSC_Profiling/Results/Segmentation/SMI/Seg/SMI_T13_F001_nuclear.csv")
colnames(nuclearseg) <- unlist(lapply(colnames(nuclearseg), function(x) {strsplit(x , split = "X0.")[[1]][2]}))
colnames(nuclearseg)[1] <- "0"
nuclearseg_raw <- nuclearseg

# Plot the nuclear cell segmentation map
nuclearseg[nuclearseg != 0] <- 0.9
display(1- Image(t(as.matrix(nuclearseg))),  method = "raster")
```

## Match cell segmentation with IF intensities 

`PanCK`, `CD45` and `CD3` immuno-flurescent intensity would be useful for 
cross-checking with computationally derived cell-type assignments downstream, 
so we will extract that now! 

```{r if_table}
frameID <- data.frame(stain = c("Membrane", "PanCK", "CD45", "CD3", "DAPI"),
           frame_id = c(1:5))
tibble(frameID)
```

In the code block below, we loop over the 3 frames that contain 
`PanCK`, `CD45` and `CD3` intensity, and every cell, to extract the IF intensity 
of the cytoplasm. It is very slow (hours for one FOV), so we will not run it. 
Instead, I am running this in parallel on the cluster (topic for another time). 

```{r if_code, eval=FALSE, include=TRUE}
if_list <- lapply(2:4, function(fid){
  print(fid)
  intensities <- data.frame(t(pbsapply(c(1:max(cellseg_raw)), simplify = T, function(cid){
    if (sum(t(cellseg_raw) == cid & t(nuclearseg) == 0) == 0) { # this is the case where the cell segmentation is pretty much just the nucleus...
      return(c(NA, NA))
    } else { 
      mean <- mean(img@.Data[,,fid][t(cellseg_raw) == cid & t(nuclearseg) == 0])
      max <- max(img@.Data[,,fid][t(cellseg_raw) == cid & t(nuclearseg) == 0])
      return(c(max, mean)) 
    }
  })))
  colnames(intensities) <- paste0(frameID[fid,]$stain, c("_max", "_mean")) 
  return(intensities)
})
if_frame <- do.call("cbind", if_list)
row.names(if_frame) <- c(1:max(cellseg_raw))
saveRDS(if_frame, file = "/Volumes/ljerby/HGSC_Profiling/Intermediate/20220410_TMA13_FOV1_intensities.rds")
```

## Computational cell-type assignments

There are some cell-types we obviously didn't stain for (e.g endothelial marker, 
fibroblast marker etc). We will use the RNA quantifications for each cell to 
assign cell-types. 

First, we load the RNA data: 

```{r rna}
tma13 <- readRDS("/Volumes/ljerby/HGSC_Profiling/Data/R_data/SMI_HGSC_TMA13_mesmer.rds")
r <- tma13$SMI_T13_F001
remove(tma13)
names(r)
```

First, we process the data by filtering out cells with too few overall counts

```{r rna_filt}
tpm <- r$tpm[,unlist(apply(r$cd, 2, sum)) > 50]
coor <- r$coor[unlist(apply(r$cd, 2, sum)) > 50,]
```

Now we use the wrappers in `opipes` to cluster the data. By "wrappers" I mean I 
wrote code that "wraps" around other code. In this case, the other code is 
`Seurat`

```{r cluster}
set.seed(123)
so <- seuratify(tpm, prj = "tma13_fov1")
so <- std_preprocess(so, nfeatures = 350)
so <- run_pca(so, n_pcs = 20, pca.assay = "RNA")
so <- umap_tsne(so, umap.flag = T, n_dims = 20)
so <- opipes::cluster(so, n_dims = 20)
```

Let's plot our clustering:

```{r plot_clust}
plot_df <- data.frame(so@reductions$umap@cell.embeddings,
                      cluster = so$RNA_snn_res.0.4)
p <- ggplot(plot_df, aes(x = UMAP_1, y = UMAP_2, col = cluster)) + 
  geom_point(size = 0.5) +
  theme_bw() + 
  coord_fixed() + 
  guides(colour = guide_legend("cluster", 
                               override.aes = list(size=3, shape = 15)))
print(p)
```


Here, we call the cell type assignments, using a cell-type assignment 
code that I've wrapped into `opipes`

```{r cell_types}
# create a list object 
q <- c()
q$tpm <- tpm 
q$genes <- row.names(tpm)
q$cells <- colnames(tpm)
q$clusters <- paste0("C", so$RNA_snn_res.0.4)

# read in your preferred cell type signatures
cell.sig <- readRDS("/Volumes/ljerby/SharedResources/Data/Signatures/cell.type.sig.rds")
cell.sig <- cell.sig$HGSC[-c(3,4,10)] # removing CD8.T, CD4.T, and cell cycle genes for now

# execute to cell-type assignment code
q <- scRNA_markers_assign(q,
                     cell.clusters = q$clusters,
                     cell.sig = cell.sig,
                     non.immune.cell.types = c('Endothelial','CAF','Malignant'),
                     immune.markers = c("CD3E","PTPRC","CD8A","CD4"))
```


What do the cell-type assignments look like in in space? 

```{r plot_df}
plot_df <- data.frame(coor, 
                      so@reductions$umap@cell.embeddings,
                      cluster = q$clusters, 
                      cell_type = q$cell.types, 
                      so@reductions$pca@cell.embeddings)
tibble(plot_df)
```

And now we use this data.frame to plot our cell-type assignments in physical space 

```{r plot}
p <- ggplot(plot_df, aes(x = x, y = y, col = cell_type)) + 
  geom_point(size = 0.5) +
  scale_color_manual(values = c("#FF5B18", "#6580f0", "#c4a360", "#bc31cc", 
                                "#3ec25f", "#fce728", "#b5b5b5")) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"), 
        panel.background = element_rect(fill = "black",
                                        colour = "black",
                                        size = 0.5, linetype = "solid"),
        panel.grid.major = element_line(size = 0.25, linetype = 'dotted'), 
        panel.grid.minor = element_line(size = 0.1, linetype = 'dotted')) +
    guides(colour = guide_legend("cluster", 
                               override.aes = list(size=3, shape = 15))) +
  coord_fixed()
print(p)
```

Isn't the comparison to the composite image uncanny? 

```{r display_comp}
display(comp, method = "raster")
```

## Running DIALOGUE for Discovering Multicellular Programs

We will now run a tool developed by Livnat, for identifying 
"multi-cellular programs". Let's do an analysis to see 
if there are multi-cullular programs between malignant, 
fibroblasts, and macrophages

The first step is to make a DIALOGUE "cell type" object
for each cell type of interest and store it in a list. 

```{r warning=FALSE}
celltypes_of_interest <-c("Malignant", "Macrophage", "CAF")
cell.types <- c()
for (celltype in celltypes_of_interest) {
  cat(paste0("\nProcessing ", celltype, "...\n"))
  idx <- which(plot_df$cell_type == celltype)
  
  cat(paste0("\tSubsetting ", celltype, " Data...\n"))
  cell_tpm <- q$tpm[,idx]
  cell_meta <- so@meta.data[idx,]
  samples <- unlist(apply(coor[idx,], 1, function(row){
                             if(row[1] < 2000  & row [2] <2000) {
                               return("niche1") 
                              } else if (row[1] >= 2000 & row[2] < 2000) {
                                return("niche2")
                              } else  if (row[1] >= 2000 & row[2] >= 2000) {
                                return("niche3") 
                              } else {
                                return("niche4")
                              }}))
  cell_q <- r$comp.reads[cell_meta$cellid]
  
  cat(paste0("\tComputing PCA for ", celltype, " Data...\n"))
  sub <- seuratify(cell_tpm)
  sub <- std_preprocess(sub, nfeatures = 350)
  sub <- run_pca(sub, n_pcs = 20, pca.assay = "RNA", verbose = F)
  
  cat(paste0("\tCompute cell type object for ", celltype,  "...\n"))
  cell.types[[celltype]] <- DIALOGUE::make.cell.type(
    name = celltype, # the cell types
    tpm = cell_tpm, # tpm 
    samples = samples, #sample 
    X = sub@reductions$pca@cell.embeddings, # pcs 
    metadata = cell_meta, #metadata 
    cellQ = cell_q # cell quality
  )
}
```

Now we can run DIALOGUE!

```{r eval=FALSE, warning=FALSE, include=TRUE}
library(DIALOGUE)
R<-DIALOGUE.run(rA =cell.types, 
                main = "20220412_Tutorial",
                k = 3, 
                results.dir = "~/Desktop/tutorial", 
                n.genes = 20)
                # conf = "cellQ",covar = c("cellQ"))
```

